---
title: "Exponential distribution simulation"
author: "Marco Tizzoni"
date: "25 March 2016"
output: pdf_document
---

# Intro
We will investigate the distribution of averages of 40 exponentials using 1,000 simulations with the following parameters:

```{r}
lambda <- 0.2
n      <- 40
nsim   <- 1000
```

The theoretical mean of the exponential distribution and the standard deviation are:
```{r}
mu <- 1/lambda
sd <- 1/lambda
```

## Show the sample mean and compare it to the theoretical mean of the distribution.
For the law of the large number, as the number of simulations increases the experimental mean approaches the theoretical mean of the distribution which should be $mu$ = `r mu`.
The following graph plots the experimental distribution mean as the number of simulations increases and the $mu$.
```{r warning=FALSE}
sim_data <- 
  data.frame( 
    nsim  = integer(0), 
    mean  = numeric(0),
    tmean = numeric(0),
    tvar  = numeric(0)
  )

for(i in 1:nsim){
  data <- rexp(n, lambda)
  
  sim_data = 
    rbind(
      sim_data, 
      data.frame(
        nsim  = i, 
        mean  = mean(data),
        tmean = mean(c(sim_data$mean, mean(data))),
        tvar  = var(c(sim_data$mean, mean(data)))
      )
    )
}

require(ggplot2)
ggplot(
  data = sim_data, 
  aes(
    x=nsim, 
    y=tmean
    )
  ) + 
  labs(
    x = "# of simulations",
    y = "mean"
    ) +
  geom_line(color='darkblue') + 
  geom_hline(aes(yintercept=mu), color='red')
```


## Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
As the distribution is approximately Normal with mean $mu$ and variance $sd^2/n$ when the number of simulation increases the variance of the distribution should approach to `r sd^2/n`. This is shown on the next graph.

```{r warning=FALSE}
ggplot(
  data = sim_data, 
  aes(x=nsim, y=tvar, colour = tvar)
  ) + 
  geom_line(color='darkblue') + 
  geom_hline(aes(yintercept=sd^2/n), color='red')
```

## Show that the distribution is approximately normal.
For the Central Limit Theorem the distribution should be approximately Normal with (mu, sd^2/n) and, in fact, the following density graph resembles that of a bell curve centered around the mean $mu$.

```{r warning=FALSE}
ggplot(
  data = sim_data, 
  aes(x=mean)
  ) + geom_histogram(aes(y=..density..),
                   binwidth=.5,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")
```