---
title: "Investigation on the distribution of averages of 40 exponentials"
author: "Marco Tizzoni"
date: "25 March 2016"
output: pdf_document
---

# Overview
We will investigate the distribution of averages of 40 exponentials with $lambda$ =0.2 using 1,000 trials.

```{r}
lambda <- 0.2
n      <- 40
nsim   <- 1000
```

The mean of the exponential distribution and the standard deviation are:
```{r}
mu <- 1/lambda
sd <- 1/lambda
```

For the Law of Large Numbers, the sample mean will converge to the theorical mean. Since the distribution will be approximated by a Normal distribution with mean $mu$ and standard deviation $sd^2/n$ we do expect the experimental mean to converge to $mu$ as the number of trials increases and the standard deviation to converge to $sd^2/n$.

## Simulations
Here we run the 1,000 simulations.
```{r warning=FALSE}
sim_data <- 
  data.frame( 
    nsim  = integer(0), 
    mean  = numeric(0),
    tmean = numeric(0),
    tvar  = numeric(0)
  )

for(i in 1:nsim){
  data <- rexp(n, lambda)
  
  sim_data = 
    rbind(
      sim_data, 
      data.frame(
        nsim  = i, 
        mean  = mean(data),
        tmean = mean(c(sim_data$mean, mean(data))),
        tvar  = var(c(sim_data$mean, mean(data)))
      )
    )
}
```

## Sample Mean versus Theoretical Mean
The following graph plots the sample mean as the number of simulation increases and the theoretical mean $mu$. As we expect as the number of simulation increases the experimental mean (blue) converges to the theoretical mean (red) of the distribution which is $mu$ = `r mu`.
```{r warning=FALSE, message=FALSE}
require(ggplot2)
ggplot(
  data = sim_data, 
  aes(
    x=nsim, 
    y=tmean
    )
  ) + 
  labs(
    title = "Sample Mean versus Theoretical Mean",
    x = "# of simulations",
    y = "Mean"
    ) +
  geom_line(color='darkblue') +
  geom_hline(aes(yintercept=sim_data$tmean[1000]), color='blue') +
  geom_hline(aes(yintercept=mu), color='red')
```


## Sample Variance versus Theoretical Variance
The same happens to the sample variance (blue): it converges to the theoretical variance $sd^2/n$ (red) when the number of simulation increases. This is shown on the next graph.

```{r warning=FALSE}
ggplot(
  data = sim_data, 
  aes(x=nsim, y=tvar)
  ) + 
  labs(
    title = "Sample Variance versus Theoretical Variance",
    x = "# of simulations",
    y = "Variance"
  ) +
  geom_line(color='darkblue') + 
  geom_hline(aes(yintercept=sim_data$tvar[1000]), color='blue') +
  geom_hline(aes(yintercept=sd^2/n), color='red')
```

## Distribution
If the distribution is approximately normal with mean $mu$ and standard deviation $sd^2/n$ we do expect we do expect a bell curve shaped density graph centered around  $mu$ = `r mu` (red). That is what we show on the next graph.

```{r warning=FALSE}
ggplot(
  data = sim_data, 
  aes(x=mean)
  ) + 
  labs(
    title = "Density function",
    x = "# of simulations",
    y = "Density"
  ) +
  geom_histogram(aes(y=..density..),
                   binwidth=.5,
                   colour="black", fill="white") +
  geom_density(alpha=.2, fill="#6666FF") +
  geom_vline(aes(xintercept=sim_data$tmean[1000]), color='blue') +
  geom_vline(aes(xintercept=mu), color='red')
```
