---
title: "Analysis of the ToothGrowth R dataset"
author: "Marco Tizzoni"
date: "25 March 2016"
output: pdf_document
---

# Overview
This report investigate the relation between different doses of Vitamin C delivered either by Ascorbid Acid or Orange Juice and the growth of teeth in guinea pigs. 
Under the assumption that the response (i.e. growth of teeth) follows a t-student distibution we use a t-test to determine whether there is a statistically significant relation between doses/delivery methods and response.

# Dataset description
In this paragraph we describe the ToothGrowth dataset, a dataset to investigate the effect of Vitamin C on tooth growth in guinea pigs.

It has three columns:

 1. *len*: the response expressed as length of odontoblasts;
 1. *supp*: delivery method, either VC (Ascorbic Acid) or Orange Juice (OJ);
 1. *dose*: three dose levels of Vitamin C (0.5, 1, and 2 mg).

# Exploratory data analysis
Figure 1 shows response for dose compared by delivery method. It looks like there is a positive relation between the subministration of ascorbic acid at doses of 0.5mg, 1.0mg and 2.0mg doses and growth of teeth.

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width=6, fig.height=3, fig.align='center'}
require(ggplot2)

ggplot(ToothGrowth, aes(factor(dose),len)) + 
  geom_boxplot( aes(fill=supp)) + 
  facet_grid(.~supp) +
  labs(
    title="Figure 1: Response for dose compared by delivery method.",
    x = "Dose",
    y = "Response"
  )
```

Figure 2 shows grooth of teeth for delivery method compared by dose: for doses of 0.5mg and 1.0mg delivery method OJ is more effective compared to VC even though at 2.0mg OJ and VC delivery methods have the same response on average.

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width=6, fig.height=3, fig.align='center'}
ggplot( ToothGrowth, aes(factor(supp), len) ) + 
  geom_boxplot( aes(fill=supp)) + 
  facet_grid(.~dose) + 
  labs(
    title="Figure 2: Response for delivery method compared by dose.",
    x = "Dose",
    y = "Response"
  )
```



## Summary of the data
In this section we provide a basic overview of the data computing mean and standard deviation for each dose and delivery method.

```{r warning=FALSE, message=FALSE, echo=FALSE}
require(dplyr)

ToothGrowth %>% 
  group_by(supp,dose) %>% 
  summarise(mean = mean(len), sd = sd(len))
```

# Key assumptions
Now we procede with our analysis under the following assumptions:

 - Variables must be independent identicaly distributed;
 - Variance of response is different when using different dose/delivery method;
 - Response follows a t-student distibution.

# Confidence intervals/hypotesis testing
In this section we perform a number of t-test using 95% confidence intervals to answer two questions:

 1. Is there a statistically significant correlation between subministration of Ascorbic Acid and growth of teeth?
 1. Which method is more effective?
 
## Ascorbic Acid is statistically correlated to growth of teeth?
H0: there is no correlation between subministration increasing doses of Ascorbic Acid and growth of teeth.

### t-test: dose 0.5 vs 1.0
Since confidence interval is (-11.983781, -6.276219) and it does not include zero **we reject H0: there is a positive correlation between doses of Ascorbic Acid and growth of teeth**.

### t-test: dose 0.5 vs 2.0
Since confidence interval is (-18.15617, -12.83383) and it does not include zero **we reject H0: there is a positive correlation between doses of Ascorbic Acid and growth of teeth**.

### t-test: dose 1.0 vs 2.0
Since confidence interval is (8.996481, -3.733519) and it does not include zero **we reject H0: there is a positive correlation between doses of Ascorbic Acid and growth of teeth**.

## Which method is more effective?
H0: VC delivery method is at least equally effective than OJ.

### t-test: dose 0.5
Since confidence interval is (1.719057, 8.780943) and it does not include zero **we reject H0: for dose levels 0.5 OJ delivery method is more effective than VC**.

### t-test: dose 1.0
Since confidence interval is (2.802148, 9.057852) and it does not include zero **we reject H0: for dose levels 0.5 OJ delivery method is more effective than VC**.

### t-test: dose 2.0
Since confidence interval is (-3.79807, 3.63807) and it does include zero **we accept H0: for dose levels 2.0 VC delivery method is as effective as VC**.


# Conclusions
 1. There is a positive correlation between the subministration of Ascorbic Acid at doses of 0.5mg, 1.0mg and 2.0mg doses and growth of teeth;
 2. at doses of 0.5mg and 1.0mg delivery method OJ is more effective compared to VC even though at 2.0mg OJ and VC delivery methods are comparable.
 
# Appendix

## Dataset description
```{r}
str(ToothGrowth)
```

## H0: there is no correlation between subministration increasing doses of Ascorbic Acid and growth of teeth.

### t-test: dose 0.5 vs 1.0
```{r}
t.test( 
  len ~ dose, 
  paired = FALSE, 
  var.equal = FALSE, 
  data = subset(ToothGrowth, dose %in% c(0.5, 1.0))
)
```

### t-test: dose 0.5 vs 2.0
```{r}
t.test( 
  len ~ dose, 
  paired = FALSE, 
  var.equal = FALSE, 
  data = subset(ToothGrowth, dose %in% c(0.5, 2.0))
)
```

### t-test: dose 1.0 vs 2.0
```{r}
t.test( 
  len ~ dose, 
  paired = FALSE, 
  var.equal = FALSE, 
  data = subset(ToothGrowth, dose %in% c(1.0, 2.0))
)
```

## H0: VC delivery method is at least equally effective than OJ.

### t-test: dose 0.5
```{r}
t.test( 
  len ~ supp, 
  paired = FALSE, 
  var.equal = FALSE, 
  data = subset(ToothGrowth, dose == 0.5)
)
```

### t-test: dose 1.0
```{r}
t.test( 
  len ~ supp, 
  paired = FALSE, 
  var.equal = FALSE, 
  data = subset(ToothGrowth, dose == 1.0)
)
```

### t-test: dose 2.0
```{r}
t.test( 
  len ~ supp, 
  paired = FALSE, 
  var.equal = FALSE, 
  data = subset(ToothGrowth, dose == 2.0)
)
```
