---
title: "Analysis of the ToothGrowth R dataset"
author: "Marco Tizzoni"
date: "25 March 2016"
output: pdf_document
---

# Overview

## Dataset description
In this paragraph we describe the ToothGrowth dataset, it is a dataset to investigate the effect of Vitamin C on tooth growth in guinea pigs.
```{r}
str(ToothGrowth)
```

It has three columns:

 1. *len*: the response expressed as length of odontoblasts
 1. *supp*: delivery method, either VC (Vitamin C) or orance Juice (OJ)
 1. *dose*: three dose levels of Vitamin C (0.5, 1, and 2 mg)

## Exploratory data analysis
```{r warning=FALSE, message=FALSE}
require(ggplot2)
ggplot( ToothGrowth, aes(factor(supp), len) ) + 
  geom_boxplot( aes(fill=supp)) + 
  facet_grid(.~dose) + 
  labs(
    title="Figure 1: Response for delivery method compared by dose.",
    x = "Dose",
    y = "Response"
  )
```

```{r}
ggplot(ToothGrowth, aes(factor(dose),len)) + 
  geom_boxplot( aes(fill=supp)) + 
  facet_grid(.~supp) +
  labs(
    title="Figure 2: Response for dose compared by delivery method.",
    x = "Dose",
    y = "Response"
  )
```

### Summary of the data
```{r warning=FALSE, message=FALSE}
require(dplyr)

summary <- ToothGrowth %>% 
  group_by(supp,dose) %>% 
  summarise(mean = mean(len), sd = sd(len))
```

## Key assumptions
 - Variables must be independent identicaly distributed;
 - Variance of response is different when using different dose/delivery method;
 - Response follows a t-student distibution.

## Confidence intervals/hypotesis testing
## Do delivery method and/or dosage affect tooth growth in guinea pigs?
