---
title: "Analysis of the ToothGrowth R dataset"
author: "Marco Tizzoni"
date: "25 March 2016"
output: pdf_document
---

# Overview
This report investigate the relation between different doses of Vitamin C delivered either by Ascorbid Acid or Orange Juice and the growth of teeth in guinea pigs. 
Under the assumption that the response (i.e. growth of teeth) follows a t-student distibution we use a t-test to determine whether there is a statistically significant relation between doses/delivery methods and response.

## Dataset description
In this paragraph we describe the ToothGrowth dataset, a dataset to investigate the effect of Vitamin C on tooth growth in guinea pigs.
```{r}
str(ToothGrowth)
```

It has three columns:

 1. *len*: the response expressed as length of odontoblasts;
 1. *supp*: delivery method, either VC (Ascorbic Acid) or Orange Juice (OJ);
 1. *dose*: three dose levels of Vitamin C (0.5, 1, and 2 mg).

## Exploratory data analysis
Figure 1 shows response for dose compared by delivery method. It looks like there is a positive relation between the subministration of ascorbic acid at doses of 0.5mg, 1.0mg and 2.0mg doses and growth of teeth.

```{r warning=FALSE, message=FALSE}
require(ggplot2)

ggplot(ToothGrowth, aes(factor(dose),len)) + 
  geom_boxplot( aes(fill=supp)) + 
  facet_grid(.~supp) +
  labs(
    title="Figure 1: Response for dose compared by delivery method.",
    x = "Dose",
    y = "Response"
  )
```


Figure 2 shows grooth of teeth for delivery method compared by dose: for doses of 0.5mg and 1.0mg delivery method OJ is more effective compared to VC even though at 2.0mg OJ and VC delivery methods have the same response on average.

```{r warning=FALSE, message=FALSE}
ggplot( ToothGrowth, aes(factor(supp), len) ) + 
  geom_boxplot( aes(fill=supp)) + 
  facet_grid(.~dose) + 
  labs(
    title="Figure 2: Response for delivery method compared by dose.",
    x = "Dose",
    y = "Response"
  )
```



### Summary of the data
In this section we provide a basic overview of the data computing mean and standard deviation for each dose and delivery method.

```{r warning=FALSE, message=FALSE}
require(dplyr)

ToothGrowth %>% 
  group_by(supp,dose) %>% 
  summarise(mean = mean(len), sd = sd(len))
```

## Key assumptions
Now we procede with our analysis under the following assumptions:
 - Variables must be independent identicaly distributed;
 - Variance of response is different when using different dose/delivery method;
 - Response follows a t-student distibution.

## Confidence intervals/hypotesis testing
## Do delivery method and/or dosage affect tooth growth in guinea pigs?
